---
title: "DATA5002 Project"
author: "HONGQING YU"
date: "2024-11-08"
output: html_document
---
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)
library(DT)
library(reshape2)
library(crosstalk)
library(dplyr)
library(here)
```

## Main factors of diabetes and how to reduce the possibility of diabetes by controlling BMI
### Heatmap of all variables in the Diabetes Data   set

Please find out which factors are the most correlated to the Outcome?

Tips: The lighter the color, the stronger the relationship.
Also, please ignore the diagonal as it represents the correlation with itself!

```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data <- read.csv("diabetes_dataset.csv", header = TRUE)
diabetes_data <- diabetes_data[diabetes_data$BMI != 0, ]
heatmap_data <- cor(diabetes_data)
ggplot(melt(heatmap_data), aes(Var1, Var2, fill = value)) +
    geom_tile() +
    theme_minimal()
```
The answer is "Glucose"!

And the second most strongly correlated variable is BMI.

Did you answer correctly?

### Boxplot of Glucose by Outcome
    Could you find out the mean, median and other features for each category?

    Tips: You can try to hovering your mouse on the chart to find the answer.
```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data$Outcome <- as.factor(diabetes_data$Outcome)
diabetes_data <- diabetes_data %>%
  mutate(if_diabetes = ifelse(Outcome == 1, "diabetes", "healthy"))

plot_ly(data = diabetes_data, 
               x = ~if_diabetes,
               y = ~Glucose,
               type = 'box',
               color = ~if_diabetes,
               colors = c("#6A9ACE","#715EA9"),
               boxmean = TRUE,
               boxpoints = 'all',  
               jitter = 0.3,       
               line = list(color = 'royalblue2'))%>%
          layout(xaxis = list(title = 'Outcome'))


```


### Density Plot of Family History by Outcome
    Do you think this plot can show if family history have a significant influence on outcome?
    Tips: You can compare two shapes and the overlapping area.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#diabetes&faminly history

ggplot(diabetes_data, aes(x = DiabetesPedigreeFunction, fill = if_diabetes)) +
  geom_density(alpha = 0.5) +
  labs(x = "Family History", y = "Density", fill = "Outcome") +
  theme_bw() +
  theme(panel.border = element_blank())

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# fig <- plot_ly(data = diabetes_data, 
#                x = ~DiabetesPedigreeFunction, 
#                color = ~Outcome, 
#                colors = c("salmon", "lightseagreen"),
#                type = 'histogram', 
#                histnorm = 'density', 
#                opacity = 0.8) %>%
#   layout(title = 'Density Plot of Family History by Outcome',
#          xaxis = list(title = 'Family History'),
#          yaxis = list(title = 'Density'),
#          barmode = 'overlay')
# 
# fig

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data$age_group <- cut(
  diabetes_data$Age,
  breaks = c(seq(21, 71, by = 10), Inf),
  right = FALSE,
  labels = c("21-31", "32-41", "42-51", "52-61", "62-71", "72-81")
)

diabetes_data$BMI_group <- cut(
  diabetes_data$BMI,
  breaks = c(-Inf, 18.5, 25, 30, 40, Inf),  
  labels = c("Under Weight", "Normal", "Overweight", "Obese", "Extremely Obese"),  
  right = FALSE  
)
diabetes_data$preg_group <- cut(
  diabetes_data$Pregnancies,
  breaks = c(0, 4, 8, 12, Inf),  
  labels = c("None-3 children", "4-7 children", "8-11 children", "More than 11 children"),  
  right = FALSE  
)

```

### Outcome Count by Age, BMI Level and Number of Children

    Which groups are you in? 

    What is the distribution of BMI levels in your group?
```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data2 <- diabetes_data %>%
  group_by(Outcome, age_group, BMI_group, preg_group) %>%
  summarise(count = n(), .groups = "drop")

diabetes_data2 <- diabetes_data2 %>%
mutate(if_diabetes = ifelse(Outcome == 1, "diabetes", "healthy"))

d <- SharedData$new(diabetes_data2)
bscols(
  widths = c(2, NA),
  list(
    filter_checkbox("age_group", "Age Group", d, ~age_group, inline = FALSE),
    filter_select("BMI_group", "BMI Group", d, ~BMI_group, multiple = FALSE)
  ),
  list(
    plot_ly(d, x = ~if_diabetes, y = ~count, type = "bar", color = ~preg_group)%>%
      layout(
        #title = "Outcome Count by Age, BMI and Number of Children",
        xaxis = list(type = "category", title = "Outcome"),
        yaxis = list(title = "Count")
      )
  )
)

```

### BMI Level Distribution and Gender
```{r echo=FALSE, message=FALSE, warning=FALSE}
obesity_data <- read.csv("obesity_level.csv", header = TRUE)
obesity_data <- obesity_data %>%
  rename(Level = X0be1dad)

#level&gender
obesity_data2 <- obesity_data %>%
  group_by(Gender, Level) %>%
  summarise(count = n(), .groups = "drop")
d1 <- SharedData$new(obesity_data2)

bscols(
  widths = c(2, NA), 
  list(
    filter_checkbox("gender", "Gender", d1, ~Gender, inline = FALSE)
  ),
  list(
    plot_ly(d1, 
            labels = ~Level, 
            values = ~count,      
            type = 'pie',
            colors = "Set1") %>%
    layout(
           #title = "Obesity Level Distribution",
           showlegend = TRUE)
  )
)
```

### BMI Level Distribution and Family History
```{r echo=FALSE, message=FALSE, warning=FALSE}

#level&Consumption of alcohol
obesity_data3 <- obesity_data %>%
  group_by(Level, CALC, family_history_with_overweight) %>%
  summarise(count = n(), .groups = "drop")
d2 <- SharedData$new(obesity_data3)

bscols(
  widths = c(2, NA), 
  list(
    filter_checkbox("family_history_with_overweight", "family history", d2, ~family_history_with_overweight, inline = FALSE),
    filter_select("CALC", "Consumption of alcohol", d2, ~CALC, multiple = FALSE)
  ),
  list(
    plot_ly(d2, 
            labels = ~Level, 
            values = ~count,      
            type = 'pie',
            colors = "Set1") %>%
    layout(
           #title = "Obesity Level Distribution",
           showlegend = TRUE)
  )
)


```

### Outcome Count by BMI Level and Trip Mode
    Could you find out the differences of trip mode between each BMI level?
```{r echo=FALSE, message=FALSE, warning=FALSE}
#level&trip mode
obesity_data3 <- obesity_data %>%
  group_by(Level, MTRANS) %>%
  summarise(count = n(), .groups = "drop")

plot_ly(obesity_data3, x = ~Level, y = ~count, color = ~MTRANS, type = "bar")

```

### Scatter Plot for Self-Comparison or Reference.
    This scatter plot shows the interval where each weight category is located distictly. 
    You can bring in your own height and weight to determine your range.
```{r echo=FALSE, message=FALSE, warning=FALSE}

fig <- plot_ly(data = obesity_data, 
               x = ~Height, 
               y = ~Weight, 
               color = ~Level, 
               colors = "Set1", 
               opacity = 0.8, 
               type = 'scatter',        
               mode = 'markers')        

fig

```



































